\clearpage
\section{Detailed Description of the Work Plan}

This section contains detailled descriptions of the project's work packages (WPs) and milestones (\cref{subsec:wp}), time and ressource planning (\cref{subsec:time}), and financial plans (\cref{subsec:finances}).

\subsection{Work packages and milestones}\label{subsec:wp}

The project comprises four following work packages: analysis, requirements, and system design (WP~1), prototype development (WP 2), prototype evaluation (WP 3), and reporting and publishing (WP 4).
Figure~\ref{fig:gantt} contains a timeline and overview of the entire project duration.

\subsubsection{WP 1 -- analysis, requirements, and system design}
%\tabelle{6}{03/26 -- 09/26}{\needcheck{X}}

%{
%\color{blue}
%Rechnung:
%
%%Wir defninieren 1 PM als eine SHK, die 40h/M arbeitet. Das müssen wir irgendwo in \cref{subsec:wp} aufschreiben, dass das die Annahme ist.
%
%Wir brauchen also 3 * 40h, 1 * 60h \textrightarrow{} 4,5 PM pro Monat über das ganze Projekt.
%
%WP1 dauert 6 Mo und läuft alleine, also \textbf{27 PM}
%
%WP2 dauert 14 Mo, davon sind zwei alleine (\textit{9 PM}), zwei sind mit WP4 geteilt (+ \textit{4,5 PM, also 4,5 * 2 / 2}), 10 sind mit WP3 und WP4 geteilt (+ \textit{15 PM, also 4,5 * 10 / 3}) == \textbf{28,5 PM}
%
%WP3 dauert 12 Mo, davon 10 mit WP3 und WP4 geteilt (\textit{15 PM}) und zwei nur mit WP4 (+ \textit{4,5 PM}) == \textbf{19,5 PM}
%
%WP4 dauert 16 Mo, davon 2 mit WP2 geteilt (\textit{4,5 PM}), und 10 mit WP2 und WP3 geteilt (+ \textit{15 PM}), und 2 alleine (+ \textit{9 PM}) == \textbf{28,5 PM}
%
%Dann musst du einmal im Text aufschreiben, dass alle gleichzeitig gleichverteilt gleichberechtigt an allem arbeiten sollen (also dass alle APs gleich viel Aufwand haben)
%
%TODO es sind nur 23 Monate aktuell, irgendwo hab ich nen Rechenfehler
%}

During the analysis and conception phase, the focus is on a comprehensive review of current publications, patents, and research reports, with particular emphasis on energy-efficient function execution, GPU orchestration strategies, and sustainable serverless platform design.
This systematic review is intended to yield sound findings that will serve as the basis for the further development of the project.

Close cooperation with the industry partner makes it possible to define the exact scope, fundamentals, and quality dimensions of the project.
Both functional and non-functional requriements are identified and recorded in a precise catalog, which includes not only technical aspects but also sustainability metrics such as energy efficiency and carbon impact.
This catalog serves as a guideline for the subsequent development and evaluation phases of the project.

Particular attention is paid to applications that can benefit from energy-efficient serverless computing and GPU-accelerated AI inference.
Identifying and considering these fields of application helps anticipate future demands on serverless platforms and ensures that the developed concepts and prototypes address both scalability and sustainability requirements in practice.
The results of this analysis directly inform the architecture of the open-source prototype to be developed in WP 2.

\subsubsection{WP 2 -- prototype development}
%\tabelle{14}{09/26 -- 11/27}{\needcheck{X}}

During the prototype development phase, the work is structured into three main parts: system design (WP 2.A), prototype implementation (WP 2.B), and the development of exemplary applications (WP 2.C).
Together, these activities ensure that the conceptual requirements defined in WP 1 are translated into a working open-source prototype that demonstrates the practical feasibility and impact of the project.

\emph{System design (WP 2.A).}
The first step is a detailed technical design of the system architecture.
Building on the catalog of requirements from WP 1, this design phase translates functional and non-functional objectives into a concrete architectural design, from which we later develop a software blueprint.
Particular emphasis is placed on mechanisms for adaptive GPU orchestration, efficient function isolation, and energy-aware scheduling.
The design process follows an iterative approach, where early prototypes are continuously validated against the requirements to ensure that every requirement is directly tied to measurable evaluation criteria such as energy efficiency, elasticity, and performance.
%Early design validation through lightweight prototyping and simulation will provide feedback loops that help refine the architecture before full implementation begins.

\emph{Prototype implementation (WP 2.B).}
%Based on the finalized design, the open-source prototype is implemented in iterative development cycles.
%The prototype builds upon established open-source technologies but is extended with novel components for sustainable serverless execution, especially so for platform components related to resource scheduling and the control plane.
%Two main subsystems form the backbone of the prototype: (1) a serverless runtime capable of managing GPU-accelerated inference tasks with minimal energy overhead, including fine-grained scaling and resource-sharing mechanisms; and (2) an orchestration layer that dynamically adapts resource allocation based on workload characteristics and sustainability metrics.
%Continuous integration and testing pipelines are established to ensure that the prototype converges toward a stable and reproducible system.
%The successful completion of this implementation marks a key project milestone.
Based on the finalized design, the open-source prototype is implemented in iterative development cycles.
The prototype builds upon established open-source technologies but introduces novel components for sustainable serverless execution, with a particular focus on GPU management and energy-aware orchestration.
Four main subsystems form the backbone of the prototype: (1) a modular serverless runtime that supports multiple programming languages and provides custom GPU abstractions through dedicated libraries; (2) a platform layer capable of handling GPU-intensive workloads and orchestrating them for maximum energy efficiency; (3) mechanisms for software- and hardware-level isolation, enabling systematic comparisons of the effects of different design choices on performance and energy trade-offs; and (4) a modular architecture that allows individual components to be swapped and extended.
Continuous integration and testing pipelines are established to guarantee that the prototype converges toward a stable, reproducible, and extensible system.

\emph{Exemplary applications (WP 2.C).}
To demonstrate the applicability and evaluate the practicality of the developed system, a set of exemplary applications is implemented.
These applications are drawn from domains where serverless AI inference has high relevance, such as computer vision workloads or natural language processing services.
Each application will later serve as a benchmark scenario to validate whether the platform fulfills the functional and non-functional requirements specified earlier.
By running these applications in controlled experiments, we can systematically assess the platform's energy efficiency, scalability, and other metrics under realistic workloads.
In addition, these applications provide reference use cases for the wider community.

%\needcheck{The successful implementation of WP 2 not only results in a functional prototype but also lays the foundation for the comprehensive evaluation activities carried out in WP 3.
%In this way, WP 2 acts as the pivotal bridge between conceptual design and empirical validation.}

\subsubsection{WP 3 -- empirical evaluation}
%\tabelle{12}{01/27 -- 01/28}{\needcheck{X}}

The empirical evaluation phase focuses on assessing the prototype developed in WP 2 against the functional and non-functional requirements defined in WP 1.
Evaluation activities are conducted iteratively throughout the development process to account for different stages of the prototype and provide continuous feedback for refinement.
The evaluation framework includes benchmarks that primarily target metrics of energy efficiency, such as GPU utilization, power consumption, and approximations overall carbon impact.
In addition, performance metrics such as latency, throughput, and scalability are measured to ensure that the system meets the expected service levels.

WP 3 also systematically compares different strategies for hardware orchestration, resource management, and isolation mechanisms at both the software and hardware levels.
This enables identification of trade-offs between energy efficiency, performance, and system overhead.
By running controlled experiments with representative AI inference workloads, the evaluation phase verifies whether the platform achieves its goals of sustainable and scalable serverless execution.
The results of WP 3 not only provide quantitative evidence of the effectiveness of the developed system but also guide potential adjustments to the prototype and inform best practices for energy-efficient serverless platform design.

\subsubsection{WP 4 -- publishing and reporting}
%\tabelle{16}{11/26 -- 03/28}{\needcheck{X}}

In the final work package, we focus on two main activities.
First, the developed software artifacts will be released as open source.
This ensures that other researchers, developers, and organizations can benefit from the findings and the sustainable serverless platform developed within the GEKO project.
The open-source release provides full access to the source code, documentation, and relevant resources, promoting transparency, reproducibility, and enabling further development and innovation within the broader community.
Preparing the software for release involves ensuring high code quality, comprehensive documentation, and proper formatting, and runs continuously alongside the ongoing development and extension of the prototype.

In the final months of the project, a comprehensive project report will be produced.
This report documents the entire project lifecycle, providing detailed insights into the analysis, development, and evaluation phases.
It offers a thorough assessment of the milestones achieved, discusses challenges encountered and the approaches taken to resolve them, and provides recommendations for potential future extensions of the platform.
The report compiles technical details, methodological approaches, results, and conclusions to give a holistic overview of the project outcomes and their significance for sustainable serverless AI computing.

\emph{Milestones.}
At the end of each work package, a clearly defined milestone marks the completion of the corresponding project phase. 
Together, these milestones provide measurable checkpoints to ensure that the project progresses according to plan and achieves its objectives.
\begin{itemize}
    \item[\textbf{M1}] WP 1 concludes with a written and elaborated catalog of requirements that serves as the foundation for all subsequent development.
    \item[\textbf{M2}] WP 2 is finalized with a working open-source prototype, demonstrating the feasibility of sustainable serverless GPU orchestration.
    \item[\textbf{M3}] WP 3 culminates in a set of realistic use cases and benchmarks that enable systematic comparison of different serverless platforms, in general, and orchestration and isolation strategies, in particular.
    \item[\textbf{M4}] Finally, WP 4 delivers a comprehensive project report documenting the entire project lifecycle, including technical results, evaluation findings, and recommendations for future work. 
\end{itemize}

\subsection{Time and ressource planning}\label{subsec:time}

%In addition, each WP description contains detailled information on its duration and required person-months (PM).
For our calculation, we define one person-month (PM) to mean a student assistant working 40 hours per month.
The work packages partly overlap to ensure smooth transitions between project stages.

\begin{figure}[!h]
    \begin{center}
        \begin{tabularx}{0.45\textwidth}{c Y c c}
            \toprule
            \textbf{WP} & \textbf{timeline} & \textbf{duration} & \textbf{PM} \\
            \midrule
            1 & 03/26--09/26 & 6 months & 27 \\
            \midrule
            2 & 09/26--11/27 & 14 months & 33 \\
            \midrule
            3 & 01/27--11/27 & 10 months & 15 \\
            \midrule
            4 & 01/27--03/28 & 14 months & 33 \\
            \bottomrule
        \end{tabularx}
    \end{center}
    \caption{
        Overview of person-months and work packages.
        The overview does not include the project lead, as it is not financed through Software Campus.
    }
\end{figure}

In total, the project requires 108 PM, which corresponds to 3 student assistants working 40 hours per month and one student assistant working 60 hours per month over the duration of the entire project.
Subtask within individual work packages are distributed equally among student assistants; hence, we calculate the PM based on the four students assistants working eqally on the current tasks, according to their monthly hours.
A more detailed timeline, which includes milestones and subcomponents of the work packages, can be found in \ref{fig:gantt}.

\subsection{Financial plan and preliminary calculations}\label{subsec:finances}

Based on the work and resource planning in the previous section, we now present the financial planning for the \name project.
This represents a preliminary calculation to the best of our knowledge and belief, based on the principle of economical and appropriate use of funds.
Overall, a total amount of 114,964.00 € is requested, with financial items for staff (student assistants), business trips (Software Campus trainings/summit), and hardware (GPUs).
An overview of all expenses can be found in \cref{tab:fp}.

\subsubsection{Staff}

As described in \cref{subsec:time}, three student assistants working 40 hours per month and one student assistant working 60 hours per month are required.
They are hired based on the TV Stud III (Tarifvertrages für studentische Beschäftigte III).
TU Berlin calculates with 17 € per hour worked by a student assistant, which covers their wage, taxes, and insurance costs.
Due to the shortage of qualified applicants in the highly competitive field of computer science, it is necessary to incentivize applications by highly qualified applicants by paying a benefit (\enquote{Zulage}) on top of the wage set by TV Stud III.
According to the establishment agreement for bonuses (\enquote{Dienstverenbarung Zulage}\footnote{\url{https://www.static.tu.berlin/fileadmin/www/10004224/Antraege/Personaldeckungszulage_sigd___.pdf}}), the bonus is set to 50 \%.
Recent hirings by the research group have shown that it is impossible to fill positions without doing so.

The technical setup for the student assistants is provided by regular funds from the research group.

\subsubsection{Business trips}

As part of the project, the project lead will attend Software Campus trainings and two Software Campus summits.
Additional travel to disseminate the results of the research conducted in this project are paid through the research group's regular funds.
Table~\ref{blub} shows an overview of all travel related to the project and their planned cost.
As all planned communication between the project lead and industry partner will be virtual, no additional travel costs emerge.

For project managers of microprojects within the Software Campus, participation in one summit per year and a total of four training sessions is required.
These events take place at various locations in Germany, which are not yet known, and will each require two overnight stays.
For project managers participating in these events, a total of 100.00~\texteuro{} is estimated for travel to and from the venue, for example by train.
In addition, based on the recommendations and rules of TU Berlin, a flat rate of 80.00~\texteuro{} per night is calculated for accommodation expenses (e.g., in hotels).

\begin{figure}[!h]
    \def\arraystretch{1.5}
    \centering
    \begin{tabularx}{\textwidth}{X l l l l l r}
        \toprule
        \bf{Event} & \bf{Dest.} & \bf{Travel} & \bf{Hotel} & \bf{Per diem} & \bf{Duration} & \bf{Total}\\
        \midrule
        Software Campus summit '26 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        Software Campus summit '27 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        SWC Training 1 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        SWC Training 2 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        SWC Training 3 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        SWC Training 4 & DE & 100.00~\texteuro & 80.00~\texteuro & 28.00~\texteuro & 3 days & 344.00~\texteuro \\
        \midrule
        $\mathbf{\Sigma}$ \bf{total expenses} &&&&&& 2,064.00~\texteuro\\
        \bottomrule
    \end{tabularx}
    \caption{Overview of planned trips and respective expenses.}
    \label{blub}
\end{figure}

\subsubsection{Hardware}

In order to successfully complete the project, additional hardware is required.
In particular, four graphics processing units will be used to sufficiently mimic real-world systems and to guarantee realistic results during the empirical evaluation phase.
The detailled hardware requirements and planned expenses are listed in table~\ref{tab:fp}.
In the appendix, we show comparisons of current prices of the GPUs.

\begin{figure}[!h]
    \def\arraystretch{1.5}
    \newcommand{\gpu}{AMD~Radeon~RX~9070~XT}
    \newcommand{\sn}{0822}
    \centering
    \begin{tabularx}{\textwidth}{l l X r}
        \toprule
        \bf{Item} & \bf{Number} & \bf{Individual item} & \bf{Expenses} \\\midrule
        Staff & \sn & Student assistant (40 hours per month) & 24,480.00~\texteuro \\
        & \sn & Student assistant (40 hours per month) & 24,480.00~\texteuro \\
        & \sn & Student assistant (40 hours per month) & 24,480.00~\texteuro \\
        & \sn & Student assistant (60 hours per month) & 36,720.00~\texteuro \\
        $\mathbf{\Sigma}$ \bf{personell expenses} &&& \bf{110,160.00~\texteuro} \\
        \midrule
        Items less than 800~\texteuro & 0831 & \gpu & 635.55~\texteuro \\
        & 0831 & \gpu & 635.55~\texteuro{}\\
        & 0831 & \gpu & 635.55~\texteuro{}\\
        & 0831 & \gpu & 635.55~\texteuro{}\\
        $\mathbf{\Sigma}$ \bf{hardware expenses} &&& \bf{2,542.2~\texteuro} \\
        \midrule
        Travel & 0846 & Software Campus & 2,064.00~\texteuro\\
        $\mathbf{\Sigma}$ \bf{travel expenses} &&& \bf{2,064.00~\texteuro} \\
        \midrule
        $\mathbf{\Sigma}$ \bf{total expenses} &&& \needcheck{\bf{114,766.20}~\texteuro} \\
        \bottomrule
    \end{tabularx}
    \caption{Overview of the financial plan.}
    \label{tab:fp}
\end{figure}





\begin{landscape}
    \thispagestyle{empty}
    \begin{figure}
        %\centering
        \hspace{-4em}
        \begin{ganttchart}[
            hgrid,
            vgrid,
            x unit=0.75cm,
            y unit title=0.75cm,
            y unit chart=0.75cm,
            bar label node/.append style={align=left, text width=7.5cm, font=\itshape},
            group label node/.append style={align=left, text width=8cm},
        ]{1}{24}

            \gantttitle{2026}{10}\gantttitle{2027}{12}\gantttitle{2028}{2} \\
            \gantttitlelist{3,...,12}{1}\gantttitlelist{1,...,12}{1}\gantttitlelist{1,...,2}{1} \\

            \ganttgroup{WP 1 -- Analysis and design}{1}{6} \\
            \ganttbar{Review of research materials}{1}{4} \\
            \ganttbar{Specification of requirements}{2}{5} \\
            \ganttbar{Exploration of possible applications}{4}{6} \\
            \ganttmilestone{M1}{6}\\

            \ganttgroup{WP 2 -- Prototype development}{7}{20} \\
            \ganttbar{System design}{7}{9} \\
            \ganttbar{Prototype implementation}{8}{20} \\
            \ganttbar{Implementation of exemplary applications}{12}{20} \\
            \ganttmilestone{M2}{20}\\

            \ganttgroup{WP 3 -- Empirical evaluation}{11}{20} \\
            \ganttbar{Development of evaluation methods}{11}{20} \\
            \ganttbar{Experiments and benchmarks}{14}{20} \\
            \ganttmilestone{M3}{20}\\

            \ganttgroup{WP 4 -- Publishing and reporting}{11}{24} \\
            \ganttbar{Open-source publication}{11}{22}\\
            \ganttbar{Project report}{20}{24}\\
            \ganttmilestone{M4}{24}
        \end{ganttchart}
    \caption{Timeline of the GEKO project.}
    \end{figure}
    \label{fig:gantt}
\end{landscape}
